Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en organization of the paper

Furthermore, we consider the problem of learning optimal actions in a causal environment. A decision maker is given the task of learning an optimal action in a causal environment in which a causal model controls the relation between actions and outcomes. We consider the case in which the causal model is unknown or partially known to the decision maker, so she has to use beliefs about possible causal models. From the decision maker point of view, having in mind a causal model allows one to ask what would have happened if another action was taken without actually performing the alternative action; therefore, being able of choosing the best action in terms of the probabilities of causing a good outcome. Also, learning a causal model of a certain environment will facilitate knowledge transfer to similar domains, as the underlying cause-effect structure has been captured and could later be used.

% añadir motivación del paper y de las teorías

\subsection{Organization of the paper}
\label{organization}
This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which use only associative information in their formulation; then, in Section \ref{causal_decision_problems} a brief introduction to Causation and Causal Graphical Models is given. In Section \ref{related_work}  we present previous approaches both to Causal Decision Making as well as its applications to interactive optimal action learning: causal reinforcement learning. In Section \ref{causal_results} we state a von Neumann-Morgenstern Theorem for Causal Decision Making that is due to Pearl, which assumes a known causal structure, as well as our main result, Theorem \ref{causal_savage} which extends Pearl's result to the case of an unknown causal structure. Finally two applications of Theorem \ref{causal_savage} are described. First, a causal analogue of Nash Equilibrium for games in which actions and consequences are causally connected is defined. Then, in Section \ref{optimal_action} we describe some optimal-action learning methods for causal environments which are able to learn an \textit{optimal intervention} under different assumptions and conditions; in particular, we describe and provide an extension for one of such works. Finally, in Section \ref{limitations} we comment on the limitations of the assumptions we consider;  and in Section \ref{conclusion} we summarize the main contributions of our work.

\section{The Classical Decision Making Theorems: von Neumann-Morgenstern and Savage}
\label{classical_theorems}
A decision problem under uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors. Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.

>>>>

Furthermore, we consider the problem of learning optimal actions in a causal environment. A decision maker is given the task of learning an optimal action in a causal environment in which a causal model controls the relation between actions and outcomes. We consider the case in which the causal model is unknown or partially known to the decision maker, so she has to use beliefs about possible causal models. From the decision maker's point of view, having in mind a causal model allows her to ask \textit{what would have happened if another action was taken} without actually performing the alternative action and thus being able to obtain the best overall action; i.e., being able of choosing the best action in terms of the probabilities of causing a desired outcome. Also, learning a causal model of a certain environment will facilitate knowledge transfer to similar domains, as the underlying cause-effect structure has been captured and could later be used.

\subsection{Organization of the paper}
\label{organization}
This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which rely on associative information in their formulation; then, in Section \ref{causal_decision_problems} we present a brief introduction to Causation and Causal Graphical Models. In Section \ref{related_work} we present previous approaches both to Causal Decision Making as well as its applications to interactive optimal action learning: causal reinforcement learning. In Section \ref{causal_results} we state a von Neumann-Morgenstern Theorem for Causal Decision Making that is due to Pearl, which assumes a known causal structure. Our main result is Theorem \ref{causal_savage} which extends Pearl's result to the case of an unknown causal structure. Finally two applications of Theorem \ref{causal_savage} are described. First, a causal analogue of Nash Equilibrium for games in which actions and consequences are causally connected is defined. Then, in Section \ref{optimal_action} we describe some optimal-action learning methods for causal environments which are able to learn an \textit{optimal intervention} under different assumptions and conditions; in particular, we describe and provide an extension for one of such works. Finally, in Section \ref{limitations} we comment on the limitations of the assumptions we consider. Finally, in Section \ref{conclusion} we summarize the main contributions of our work.

\section{The Classical Decision Making Theorems: von Neumann-Morgenstern and Savage}
\label{classical_theorems}
A Decision Droblem under Uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors \citep{bernardo2000bayesian,gilboa2009decision}. Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.
